Developing an automated iterative near-term forecasting system for an ecological

### 2018-06-29 manuscript/response_to_review.Rmd

---
title: "Response to Review"
output: pdf_document
---

## Associate Editor

**The main issue regards the motivation (reviewer 1) and target audience (reviewer 2) of the application. If this is meant for researchers that are proficient at modeling and coding, it should be made clearer how they benefit from this automated application (wouldn't they already have similar protocols and pipelines in place in their routine research?). If instead, this is for empiricists, or environmental managers, with little coding experience, the manuscript is too technical and the application perhaps should be made more 'user-friendly' (reviewer 1). My suspicion (and reviewer 2's) is that it aims for that increasingly common middle person (the computationally literate ecologist). However, as Reviewer 2 points out, it is not clear how such audience should interact with the defined pipeline and how, for example, they could propose new models. A fourth (very attractive) possibility is that this is meant as an interface for people with different skills and motivations to collaborate (e.g. to foster the feedback between scientific development and applied use). If this is the case, the protocols of interaction need to be clarified, as well as the audience for this article (e.g. is it the computational ecologist who will read this article and propose applied scientists to work with her/him through this system?). Defining the audience clearly should guide how to improve the application and manuscript in a way that it ensures its use and long-term survival as an important short-term forecasting tool.**

We have expanded the last paragraph of the Introduction to address the question of target audience and split the opening paragraph of the Introduction into two paragraphs to provide expanded treatment of the motivation for this approach including the addition of three examples.

The AE and reviewer are correct that we have target this paper primarily at the "increasingly common middle person" and "researchers that are proficient at modeling and coding". Thanks to the expanded efforts at computational training we believe that many researchers have the necessary skills to implement these sorts of approaches and we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.

That said, we certainly recognize that not all ecologists and evolutionary biologists have the necessary skills to implement this sort of pipeline (just as many do not have the skills to implement a Bayesian statistical analysis). As a result, a single paper in this space won't be perfect for everyone. We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate. To reach researchers with relatively basic computational skills we have recently completed a paper focused on the data management component of this pipeline that focuses on providing a ground up introduction to this kind of automation in an ecological context (https://doi.org/10.1101/344804). We hope that the combination of these two papers will successfully introduce this general class of approaches and concepts to a broad array of researchers.

We disagree with the assumption that researchers that are proficient at modeling and coding would already have similar protocols and pipelines in place in their routine research?)". We are active in this field and aware of very few efforts of this scope in ecology. These researchers will have the foundational tools to develop similar systems (or contribute to this one).

# Reviewer: 1

**Overall this is a clear (and very thorough) description of a useful set of methods and tools for addressing a compelling need in ecology.  In this age of increased skepticism of science among the general public and increased awareness among scientists of the mischief that statistics can cause, this kind of automatically updating system seems like the most honest and open approach to science possible.**

Thanks!

**My only real comment/question regards the motivation for the automatic and iterative nature of this. The full data processing to forecasting pipeline is great because it integrates data processing, model fitting and model comparison.  But I suppose the skeptical reader might wonder, can’t all of this be done without automatic and iterative part? Can’t we compare model predictions the old fashion way, using held out data, not by forecasting truly novel data? The response might be that forecasts themselves are inherently useful for planning.  Again the skeptic might question whether there really is a demand for such short-term forecasts? Maybe more examples could be given. My hunch is that the real benefit of iteratively adding truly novel data is that it makes the science as honest as possible. I guess I’m asking that the text somewhere address the skeptic’s questions above.**

There is definitely a need and demand for short-term forecasting. This was laid out in detail in Dietze et al. (2018; PNAS), but we agree that the current paper benefits from providing more direct treatment of this. We have significantly expanded the opening paragraphs of the introduction to provide specific examples and further emphasize the value of this approach.

**I might also suggest differentiating parameter estimation and model development. Forecasts can improve by adding new models, or by adding new data which then improves parameter estimation in the old models.  The automatic pipeline is great at the second part of this, continually adding data and re-fitting the old models to the new data and then producing new forecasts.  However, since there was so much data in the Portal project to begin with, I doubt that parameter estimates in the models have changed much since the automatic pipeline was established (maybe they have?).  However, in a new project that was just getting started (e.g. NEON), the main improvements in forecast skill early on would be in adding more data.  Eventually, however, the models might stop improving and reach some prediction limit imposed by the structure of the model.  As the authors point out in the discussion, finding this point will be key for improving ecological theory. I think a lot of ecologists think their models are useful, they just need better parameter estimates.  My sense is that actually, most ecological models are just crap for forecasting, it’s not for lack of data! Maybe this is an obvious point, but the role of improved parameter estimation in this whole endeavor kind of gets lost for me.  Does the pipeline even save the fitted model parameters? Would that be worthwhile?**

There are a number of interesting points in this paragraph that we feel are beyond the scope of this current paper. We have already described the importance of new models and how we have facilitated adding them. We have now further expanded this section to describe how those external to the group can contribute improved models. Introducing and discussing the different implications of differnt forms of error and how the are associated with increasing in data vs. improvements in models is an interesting subject for forecasting, but is out of scope of this paper. It would definitely be something interesting to explore in future work.

**In Figure 1, I feel that something is missing here about how other scientists are supposed to interact with the data and forecasts.  The manuscript makes it pretty clear that this is not just for internal use, but is it actually inviting other labs to add models and make predictions? How specifically would another lab contribute a model to the project? Would they fork the project on github, develop a new model and then do a pull request? Or would they create a separate prediction archive themselves? Maybe add an arrow in Fig 1. showing how new models can be plugged into the forecasting work flow.  Box 1 addresses some of this but I guess I want more details.**

We have added description of how others can contribute to the Modeling and Forecasting section of the paper. We tried to also add this to the figure, but could not find a way to do so without making the figure overly crowded and complicated.

**Some other miscellaneous questions:**
**Can models be dropped from the ensemble?**

This is not currently done at the moment, but is on the development roadmap.

**What if old data needs to be corrected or updated? That could change everything, including the interpretation of the archived predictions. Is there a way to handle that?**

This is a great question about a tricky problem in forecast archiving that we have encountered and are still trying to determine out the optimal way to address. One can update the previous forecast files and use the versioning to maintain a copy of the forecasts that contained the underlying error or leave them as they are an store supplementary data indicating that forecasts prior to the selection should not be used in later analyses. Since we aren't happy with any of the current solutions, and this is a fairly detailed point related to archiving, we think it is best to leave this for later discussion.

**Minor comments:**

**Lines 63 to 83:  Could this section include some examples that demonstrate the value of iterative near-term forecasting from other fields? Could it be argued that iterative near term forecasting has led to progress in other fields? Say meteorology and political science?**

We have now mentioned the established benefits for iterative near term forecasting in meteorology and added several citations for readers interested in exploring this further.

**Figure 1: Another point on figure one.  What about model improvement? It seems like there should be a feedback loop somewhere showing how models get refined, added, dropped etc.**

We tried to add a visual representation of this to Figure 1, but could not make it work without making the figure overly crowded. We have added some additional discussion of this "Modeling and Forecasting" section of the paper.

**There’s a lot of technical terms and software names.  Worth adding a glossary or table listing all the tools and acronyms used in the paper.  Github, Travis, CI, QA/QC, Docker, Rocker, Software containers, portalData, portalr, portalPredictions, Zenodo, data latency, etc.**

We have added a glossary.

**Figure labels are doubled.**

Fixed.

**Travis and github icons are very small in Fig. 1**

We have increased the size of the icons.

**comment on the website:  portal.naturecast.org/evaluation.html  – looks like there’s a word missing in the caption explaining the coverage graphs.  “Coverage: ... equal to . “**

Fixed. Thanks.

**comment on the website: http://portal.naturecast.org/index.html  – Figures need more information/captions.  Which model is producing the blue lines (the forecasts).  Is this the ensemble model?**

Yes. We have clarified this on the website.

**Line 163:  This section here about dependencies seems a little out of place, or at least the transition from the CI section to the details of dependencies management seems abrupt.**

There are two main components to Continuous Analysis as defined by Beaulieu-Jones & Greene (2017). We have one paragraph on each in this section on Continous Analysis. We're not sure how else to handle this, but if the reviewer has any specific suggestions we're happy to consider them.

**Figure 2:  The figure is pixelated and hard to read (might just be the draft version I have).**

The raw images are all high resolution so presumably this will be fixed in production if the paper is accepted.

## Reviewer: 2

**I enjoyed reading the manuscript by White et al, which provides a very clear description of using cutting-edge software development technologies for ecology by automating near-term forecasting. The manuscript is well thought-out, and well-written and I think should be accepted for publication. I have a few minor line-by-line suggestions and a broader question/theme that was in my mind throughout reading the article that I would suggest the authors consider.**

Thanks!

**Minor line-by-line suggestions:**

**L48: Could you cite a few here, maybe give a sentence or two overview of what they do/what questions they seek to answer (are we talking about continuous temperature sensors, camera traps, terrestrial/aquatic/marine?)**

We have split the first paragraph into two to allow us to add multiple examples and associated citations.

**L102: Like L48 above, please provide citations and examples with this statement**

We have added examples and citations to this sentence.

**Fig 1:**

**- The first time the reader sees "GitHub" and "Travis": please add citations and descriptions (and/or reshuffle when Fig 1 is cited within the text so you can describe it first in the text)**

There is a challenging order issue here. We want readers to be able to see the stages of the forecasting pipeline before delving into the details. This is Figure 1a, which is referenced in the text before the definitions of Travis & GitHub. That section of the figure/legend does not reference those terms. Travis was cited and described before Figure 1b was referenced. We have updated the text so that this is also true for GitHub. We feel that this is the best compromise in that it keeps the two directly related pieces of Figure 1 together, allows the reader to see the big picture of the pipeline before diving into the details, and introduces the associated terms before the relevant section of the figure is referenced in the text.

**- Consistency: the legend describes the Travis icon but not the GitHub icon**

Added.

**- A, B are not labeled in the figure**

Fixed.

**This paper is written to quite a narrow audience: ecologists that already code and could benefit from this kind of automation. This is fine. But many researchers do not yet have the coding skills to benefit from automation (the audiences for Software Carpentry, who the authors mention). To me, it seems that in ecology there are two computational camps at the moment: ecologists who are ready for the workflows mentioned in this paper, and those beginning to engage through Software Carpentry. This is obviously over-simplified, but as I read this manuscript I found myself thinking that it would be nice to try to speak to both camps, instead of just the one with the skills for this automation. How do people get there? How do we get ecologists to the point where they too can develop a workflow like the one in this manuscript?**

A single paper in this space won't be perfect for everyone. We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate. Thankfully given recently expanded efforts to provide more advanced computational training to scientists we disagree that it is currently written for "quite a narrow audience". That said, we agree with the reviewer that reaching out to researchers with relatively basic computational skills is important. To address this broader need for introductory material on automation and continuous integration in ecology, we have recently completed a paper focused on the management of data that is regularly updated, which focuses on communicating at an introductory level, including the inclusion of a tutorial for helping beginners to this area get started (https://doi.org/10.1101/344804). We hope that the combination of these two papers will successfully introduce these approaches and concepts to a broad array of researchers.

**I was happy to see the discussion of team science (~L421), but training and skill-building immediately came to mind, and I didn't know who the "team" was, it seems that it is the Weecology team, not including the Portal Project. Can the Portal Project maintain this on their own? I still didn't have a sense whether Florida researchers created something for Arizona field scientists that will be maintained by the Florida scientists, or whether the Arizona group was trained to maintain it on their own. My assumption is that the Arizona team would be in the Software Carpentry camp, and I'm thinking about how to bridge these camps and empower more scientists with the skills to be able to understand the integrative process the manuscript describes. The goal is to get all ecologists into the automation camp, how do we do it?**

**I would suggest adding some sentences that address this to the intro and discussion. This could traditionally be seen as out of the scope of the manuscript, but that will continue to perpetuate the two camps instead of bringing the Software Carpentry camp towards the automation camp.**

As is already described in the paper, the field researchers are an active part of the team that develops and runs this system. Both components of the team are part of Weecology. As we've mentioned above the clear distinction between empiricists and computional folks really isn't valid anymore and the field researchers involved in this project all have strong computational skills. We have modified the language in this section to clarify how the team works. While we agree that "The goal is to get all ecologists into the automation camp", and we are working hard on multiple fronts to make this a reality, this is a larger challenge that is beyond the scope of this current paper.

### 2018-06-29 manuscript/portal_iterative_forecasting_ms.Rmd
Improvements to ms Figure 1 and legend
Response to reviews
* Adds addition examples and justification
* Addresses confusion about target audience
* Makes a number of minor clarifications and improvements
* Responds to reviewers
* Switches gif image to png to improve useability with most latex installs

Co-authored-by: Morgan Ernest <morgan@weecology.org>

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology. Since then, an increasing number of ecological forecasts are being published. Most of these forecasts, however, are made once, published, and never assessed or updated. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build. Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petchey2015; @dietze2018]. More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

@@

While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement, and few of their recommendations are in widespread use in ecology today. We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going, long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008]. We constructed an automated forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis. In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.

@@

![Figure 1. a) Stages of the forecasting pipeline. To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure. The stages are interdependent, with outputs from one stage forming the inputs for the subsequent stage. Tasks in all stages are run using code written in R. b) Continuous integration system. Each box denotes the core infrastructure used for each stage of the forecasting pipeline. Continuous integration (denoted by the Travis icon, a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box). Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box). ](figs/Figure1.png)

@@

A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline. We employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components. Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI). CI combines computing environments for running code with monitoring systems to identify changes in data or code. Essentially, CI is a computer helper who watches the pipeline and, when it sees a change in the code or data, runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end. This is useful for iterative near-term forecasting because it does not rely on humans to create new forecasts whenever new models or data are added. These tools are common in the area of software development, where they are used to automate software testing and integrate work by multiple developers working on the same code base. However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017]. Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time). Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist [@beaulieu2017]. We use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).

@@

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b). We use a plugin infrastructure to allow new models to be easily added to the system. This approach treats each model as an interchangable black box; all models have access to the same input data and generate the same structure for model outputs (Figure 2). During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future. As long as a model script can load the provided data and produce the appropriate output, it will be run and its results incorporated into the rest of the forecasting system.

![Figure 2. Demonstration of plugin infrastructure. All model scripts (represented here by the example AR1.R) are housed in a single folder. Each model script uses data provided by the core forecasting code (represented here by rodent.csv) and returns its forecast outputs in a predefined structure that is consistent across models (represented here by the example 2017_12_08forecasts.csv). Outputs from all models run on a particular date are combined into the same file (i.e. 2017_12_08forecasts.csv) to allow cross-model evaluations. Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.gif)

@@

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting [forecasting on already collected data; @jolliffe2003] to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12-month period can immediately be used to evaluate the model. Since hindcasting is conducted using data that has already been collected, it allows model comparisons to be conducted on large numbers of hindcasts and provides insight into which models make the best forecasts without needing to wait for new data to be collected [@harris2018]. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.

@@

![Figure 3. Screen capture of the homepage of the Portal Forecasting website (http://portal.naturecast.org). This site contains information on the most current forecasts, evaluation of forecast performance, and general information about the species being forecast.](figs/website_fig.png)

@@

Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise. It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab. When teams have a breadth of expertise, communication can be challenging [@winowiecki2011]. We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group. The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists  (http://weecology.org). Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts. In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions. This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions. Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs. This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves. To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2016].

@@

>>>>

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology. Since then, an increasing number of ecological forecasts are being published that focus on a societally important questions from daily to decadal time scales [@dietze2018]. At daily scales, ecological forecasts predict the occurance of environmental issues like toxic algal blooms [@stumpf2009]. At quarterly scales forecasts are used to predict the stocks of fisheries and make decisions about quotas for fishing [@NOAA2016]. At decadal time scales, ecological forecasts are used to predict how biodiversity will change as it responds to anthropogenic influences [@harris2018]. These forecasting examples highlight the important role that ecological forecasts play in recasting ecological knowledge in societally relevant ways and also improve our understanding of ecological systems by testing the ability of our models to predict how systems will change in the future [@dietze2018, @harris2018].

While some of the examples given above (e.g., fisheries stock estimates) are regularly repeated, most ecological forecasts are made once, published, and never assessed or updated [@dietze2018]. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build. Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petchey2015; @dietze2018]. More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general. This approach has helped accelerate forecasting ability in other fields including meteorology [@kalnay2003, @mcgill2012, @bauer2015]. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

@@

While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement (e.g., making quality data available in near real-time and automatically rerunning forecasts in reproducible ways), and few of their recommendations are in widespread use in ecology today [e.g., @wilson2014; @stodden2014; @yenni2018]. We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going, long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008]. We constructed an automated forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis. In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems. For those interested in implementing iterative forecasting, either on their own or as part of a team, this paper will provide a roadmap for how to build such a system and what skills will be helpful to do so. For readers looking for an introduction to automation and continous integration in an ecological context, we recommend our paper on data management for continuously collected data, which includes a tutorial on how to set up some of the aspects of automation described in this paper (@yenni2018).

@@

![a) Stages of the forecasting pipeline. To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure. The stages are interdependent, with outputs from one stage forming the inputs for the subsequent stage. Tasks in all stages are run using code written in R. b) Continuous integration system. Each box denotes the core infrastructure used for each stage of the forecasting pipeline. Continuous integration (denoted by the Travis icon, a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box). Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box). The code for driving different stages of this pipeline is stored on GitHub (denoted by the GitHub icon, an "octocat").](figs/Figure1.png)

@@

A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline. We employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components. Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI). CI combines computing environments for running code with monitoring systems to identify changes in data or code. Essentially, CI is a computer helper who watches the pipeline and, when it sees a change in the code or data, runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end. This is useful for iterative near-term forecasting because it does not rely on humans to create new forecasts whenever new models or data are added. These tools are common in the area of software development, where they are used to automate software testing and integrate work by multiple developers working on the same code base. However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017]. Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time). This continuous integration integrates directly with GitHub (https://github.com), the online repository where we store the associated code and data. Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist [@beaulieu2017]. We use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).

@@

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b). We use a plugin infrastructure to allow new models to be easily added to the system. This approach treats each model as an interchangable black box; all models have access to the same input data and generate the same structure for model outputs (Figure 2). During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future. As long as a model script can load the provided data and produce the appropriate output, it will be run and its results incorporated into the rest of the forecasting system. This means that anyone can add a new model to the existing system by: 1) creating their own copy of the project (typically by forking the project on GitHub); 2) developing a new model; and 3) submitting a pull request to our repository.

![Demonstration of plugin infrastructure. All model scripts (represented here by the example AR1.R) are housed in a single folder. Each model script uses data provided by the core forecasting code (represented here by rodent.csv) and returns its forecast outputs in a predefined structure that is consistent across models (represented here by the example 2017_12_08forecasts.csv). Outputs from all models run on a particular date are combined into the same file (i.e. 2017_12_08forecasts.csv) to allow cross-model evaluations. Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.png)

@@

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting [forecasting on already collected data; @jolliffe2003] to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12-month period can immediately be used to evaluate the model. Since hindcasting is conducted using data that has already been collected, it allows model comparisons to be conducted on large numbers of hindcasts and provides insight into which models make the best forecasts without needing to wait for new data to be collected [@harris2018]. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate. As the performance of different models is understood through evaluation of forecasts and hindcasts models can be refined or removed from the system or ensemble to iteratively improve the resulting forecasts.

@@

![Screen capture of the homepage of the Portal Forecasting website (http://portal.naturecast.org). This site contains information on the most current forecasts, evaluation of forecast performance, and general information about the species being forecast.](figs/website_fig.png)

@@

Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise. It is rare to find such breadth within a single individual, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab. When teams have a breadth of expertise, communication can be challenging [@winowiecki2011]. We found a shared base of knowledge related to both the field research and computational skills was important for the success of the group. The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists (http://weecology.org). Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and lab training efforts. In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions. This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions. All members of the team actively participated in the development of the forecasting pipeline. Given the current state of tools for forecasting, forecasting teams require some experience in working with continuous integration and APIs. This means either interdisciplinary teams or additional training will often be required for creating these pipelines until tool development improves. To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future [@read2016].

@@

# Box 2. Glossary of terms

**CI.** 'Continuous Integration.' The practice of continuouslly building and testing a code base as it is developed. 
**Data latency.** The time it takes for data to be available for use.
**Docker.** An open-source Linux program for containerization (see software container).
**git.** An open-source version control system.
**GitHub.** A web-based host for git projects. Other options for a similar service include GitLab or Bitbucket.
**PortalData.** The git repository for the Portal data, found on GitHub.
**portalPredictions.** The git repository for the forecasts made using Portal data, found on GitHub.
**portalr.** An R package for using the Portal data.
**QA/QC.** 'Quality Assurance.' Testing the quality of a product. 'Quality Control.' The process of ensuring the quality of a product.
**Rocker.** A project making it easy to use Docker containers in the R environment.
**Software container.** Allows a developer to package up an application with all of the parts it needs to run reliably.
**testthat.** R package used to set up automated testing for QA/QC.
**Travis.** A continuous integration service that integrates easily with GitHub and R. Examples of similar programs are Jenkins or CodeShip.
**Unit test.** A component of quality control in which each smallest testable part of software is formally tested.
**Zenodo.** An open data archive that integrates easily with GitHub.
